---
title: "R Notebook"
output: html_notebook
---
## Loading in libraries

```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(janitor)
library(readr)
library(here)
```

To do:
- Assertive programming
- Unit tests
- Analysis
- Doing readme
- Changing to public on git

## Brief introduction to the dataset
This is a dataset with results from two decathalon competitions, Decastar and
Olympics.

## List of assumptions
- That the names were not needed.

## Steps to clean the code
- Moving the names from the rownames, and into a column for names.
- Deleting the row names.
- Moving the row names column to the front of the dataset.
- Cleaning all the column names using janitor.
- Manually cleaning the remaining column names, by manually renaming them.
- Pivoting the table.
- Saved to csv.

Reading in cleaned data
```{r}
# reading in cleaned data
cleaned_decathalon_data <- read_csv("clean_data/decathalon_clean.csv")

<- read_csv(here::here())


```

## Answers to questions from the brief

Who had the longest long jump seen in the data?
```{r}
select names and long_jump
filter by highest

```

What was the average 100m time in each competition?
```{r}
select decastar and 100m
mean
find name

same for olympics




```

Who had the highest total points across both competitions?
```{r}
select decastar
highest of the points
who


select olympics
highest of the points
who


```

What was the shot-put scores for the top three competitors in each competition?
```{r}
select rank
filter shot put only
group by competitor and rank >= 1
see point scores

```
What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?
```{r}
filter competitors where 400m > 50
mean points

filter competitors where 400m < 50
mean points

```

## Other interesting analyses and conclusions
```{r}

```

