---
title: "R Notebook"
output: html_notebook
---
## Loading in libraries

```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(janitor)
library(readr)
library(here)
```

## Brief introduction to the dataset
This is a dataset with results from two decathalon competitions, the Decastar 
which is an annual French competition, organised by World Athletics,
and the Olympics which is organised by the IOC. 

## List of assumptions
- That the names were not needed.

## Steps to clean the code
- Moving the names from the row names, and into a column for names.
- Deleting the row names.
- Moving the row names column to the front of the dataset.
- Cleaning all the column names using janitor.
- Manually cleaning the remaining column names, by manually renaming them.
- Pivoting the table.
- Saved to csv.

### Note
- Unit tests and assertive programming are outstanding. 

Reading in cleaned data
```{r}
# reading in cleaned data
cleaned_decathalon_data <-
  read_csv(here::here("~/GitHub/dirty_data/task_1/dirty_data_codeclan_project_louise_task1/clean_data/decathalon_clean.csv"))
```

## Answers to questions from the brief

Who had the longest long jump seen in the data?
```{r}
cleaned_decathalon_data %>%
  select(long_jump, olympic, decastar) %>%
  arrange(desc(long_jump)) %>%
  slice_max(1)
# finding the longest long jumper 
```

What was the average 100m time in each competition?
```{r}
cleaned_decathalon_data %>%
  select(hundred_metre, decastar) %>%
  summarise(mean_decast_hundred = mean(hundred_metre))
#finding the mean for the Decastar hundred_metre

cleaned_decathalon_data %>%
  select(hundred_metre, olympic) %>%
  summarise(mean_olymp_hundred = mean(hundred_metre))
#finding the mean for the Olympic hundred_metre

```

Who had the highest total points across both competitions?
```{r}
cleaned_decathalon_data %>%
  select(points, decastar) %>%
  filter(!is.na(decastar)) %>%
  arrange(desc(points)) %>%
  slice_max(1)
# Finding the competitor with the highest point total in the Decastar Decathalon

cleaned_decathalon_data %>%
  select(points, olympic) %>%
  filter(!is.na(olympic)) %>%
  arrange(desc(points)) %>%
  slice_max(1)
# Finding the competitor with the highest point total in the Olympics Decathalon
```

What was the shot-put scores for the top three competitors in each competition?
```{r}
cleaned_decathalon_data %>%
  select(shot_put, decastar) %>%
  filter(!is.na(decastar)) %>%
  arrange(desc(shot_put)) %>%
  slice(1:3)
# finding the top three shot put scores for Decastar Decathalon

cleaned_decathalon_data %>%
  select(shot_put, olympic) %>%
  filter(!is.na(olympic)) %>%
  arrange(desc(shot_put)) %>%
  slice(1:3)
# finding the top three shot put scores for Olympic Decathalon
```
What was the average points for competitors who ran the 400m in less than 50
seconds vs. those than ran 400m in more than 50 seconds?
```{r}
cleaned_decathalon_data %>%
  select(four_hundred_metre, points) %>%
  filter(four_hundred_metre <= 50) %>%
  summarise(mean(points))
# finding the average point total for competitors who ran the four hundred metre
# in less than or equal to 50s. 

cleaned_decathalon_data %>%
  select(four_hundred_metre, points) %>%
  filter(four_hundred_metre > 50) %>%
  summarise(mean(points))
# finding the average point total for competitors who ran the four hundred metre
# in more than 50s.
```

## Other interesting analyses and conclusions
```{r}

```

